\chapter{KUnit - Kernel Unit Test Framework}

KUnit is a PUnit based portable unit testing framework that is modified to allow
running unit tests in kernel space, called in-kernel unit testing. This is done
because it's not easily possible to compile kernel code to user space to link
against with unit test code that could be run in user space.

When kunit is compiled with the kernel tests are exposed in sysctl MIB tree
under debug.test. Any test can be invoked by writing a value other than zero
to the corresponding variable in MIB. While test is executing the user thread
is blocked until test tests have been run. Test results are written at
"realtime" to the kernel logger.

KUnit and unit tests are built into the kernel if \verb+configKUNIT+ flag is set
in config file. This creates a \verb+debug.test+ tree in sysctl \ac{MIB}. Unit
tests are executed by writing value other than zero to the corresponding
variable name in MIB. Tests results are then written to the kernel log.
Running thread is blocked until requested test has returned.

\section{Assertions}

\begin{itemize}
\item \verb+ku_assert(message, test)+ - Checks if boolean value of test is true
\item \verb+ku_assert_equal(message, left, right)+ - Checks if
      \verb+left == right+ is true
\item \verb+ku_assert_ptr_equal(message, left, right)+ - Checks if left and
      right pointers are equal
\item \verb+ku_assert_str_equal(message, left, right)+ - Checks if left and
      right strings are equal (strcmp)
\item \verb+ku_assert_array_equal(message, left, right, size)+ - Asserts that
      each integer element i of two arrays are equal (\verb+strcmp()+)
\item \verb+ku_assert_str_array_equal(message, left, right, size)+ - Asserts
      that each string element i of two arrays are equal (\verb+==+)
\item \verb+ku_assert_null(message, ptr)+ - Asserts that a pointer is null
\item \verb+ku_assert_not_null(message, ptr)+ - Asserts that a pointer isn't
      null
\item \verb+ku_assert_fail(message)+ - Always fails
\end{itemize}
