\part{File System}

\chapter{File System Abstraction}

\section{General Principles}

During this part of the document terms \acs{vnode} and \acs{inode} are
sometimes used interchangeably. Historically inode was a way to index files
in Unix-style file systems.\cite{Wikipedia:inode} While Zeke continues this
fashion it adds a concept of vnodes that are used as an abstraction level
between actual file systems and \acf{vfs}. This is done mostly in the same way
as on most of modern Unices today.

In Zeke vnodes are always used as a primary access method to the files in
a file system, inodes are only accessed internally within the file system
implementation code if the file system supports/uses inodes. Expected behavior
is that a vnode and a vnode number exist in the system as long as some process
owns a reference to that vnode. After there is no more references to a vnode
it's not guaranteed that the vnode exist in the memory and it cannot be
retrieved by using its vnode number. In fact normally vnodes can't be accessed
by their vnode number but a vnode number is guaranteed to be unique within
a single file system, superblock + vnode, but same vnode might be in use within
another file system and the number can be also reused.

To simplifify understanding how \acs{vfs} works in Zeke it can be described as
an object storage where file objects are first searched, found and the
associated (open) with a process. When a file is opened the process owns a
pointer to the file descriptor that contains some state information and pointer
to the actual file vnode. The vnode of the file is itself an object that knows
where contents of the file is stored (physical file system, superblock pointer)
and who knows how to manipulate the data (pointer to the vnode operations struct).
In fact vnode number itself is pretty much redundant and legacy information for
Zeke that is only provided for compatibility reasons, the actual access method
is always by a pointer reference to an object.

\section{Kernel Interface}

Kernel interface to the actual file system drivers and file system superblocks
is built around virtual function structs defined in \acs{vfs} header file
\verb+fs.h+ and some user space header files defining unified data types.

A new file system is first registered to the kernel by passing a pointer to
fs struct that is a complete interface for mounting a new superblock and
interacting with the file system (note the difference between a file system
(driver) and a file system superblock that is referencing to the actual data
storage, while fs driver is accessing the superblock. The file system struct
is shown in \ref{list:fs}.

When superblock is mounted a superblock struct pointer is returned. This pointer
servers as the main interface to the newly mounted file system. Superblock is
defined as in listing \ref{list:fs_sb}. By using superblock function calls it's
possible to get direct references to vnodes and \acs{vnode} operations e.g. for
modifying file contents or adding new hard links to a directory node.

\lstinputlisting[label=list:fs,caption=fs struct definition.]{fs/fs.c}
\lstinputlisting[label=list:fs_sb,caption=superblock struct definition.]{fs/fs_superblock.c}

\section{VFS hash}

\verb+vfs_hash+ is an optional hashmap used for vnode caching for physically
slow file systems. VFS itself doesn't use this caching for any purpose so
it is completely optional to insert anything into there.

\chapter{fatfs}
\input{fs/fatfs}

\chapter{procfs}
\input{fs/procfs}

\chapter{ramfs}
\input{fs/ramfs}

\chapter{devfs}
\input{fs/devfs}
