\chapter{sysctl}

The Zeke sysctl mechanism uses hierarchically organized \ac{MIB} tree as a
debugging and online configuration interface to the kernel. This is extremely
useful for example when testing scheduling parameters. Instead of recompiling
after every parameter change it is possible to change kernel's internal
parameters at run time by using sysctl interface.

There is only one syscall for sysctl which handles both reading/writing a
\ac{MIB} variable and queries to the MIB.

\section{Magic names}

There is some magic OID's that begins with \verb+{0,...}+ that are used for
queries and other special purposes. Particularly all OID's begin with 0 are
magic names. Currently allocated magic names are described in table
\ref{table:sysctlmagic}.

\begin{table}
\caption{sysctl magic names.}
\label{table:sysctlmagic}
\begin{tabular}{lll}
Name                & Internal function        & Purpose\\
\hline
\verb+{0,1,<iname>}+ & \verb+sysctl_sysctl_name()+     & Get the name of a MIB variable.\\
\verb+{0,2,<iname>}+ & \verb+sysctl_sysctl_next()+     & Get the next variable from MIB tree.\\
\verb+{0,3}+            & \verb+sysctl_sysctl_name2oid()+ & String name to integer name of the variable.\\
\verb+{0,4,<iname>}+ & \verb+sysctl_sysctl_oidfmt()+   & Get format and type of a MIB variable.\\
\verb+{0,5,<iname>}+ & \verb+sysctl_sysctl_oiddescr()+ & Get description string of a MIB variable.
\end{tabular}
\end{table}

\section{Adding new sysctl entries}

\subparagraph{Nodes}
New nodes containing sub-entries can be created with \verb+SYSCTL_NODE+ macro
like shown in listing \ref{list:sysctl_node}.

\lstinputlisting[label=list:sysctl_node,caption=Sysctl node macro.]{sys/sysctl_node.c}

In order to populate variables and nodes under newly created node the node
should be declared with \verb+SYSCTL_DECL(<name>);+, this can be done either in
\verb+sysctl.h+, in the header file of the subsystem/mode or locally in the
source code file. In the latter case the new node won't be available in
global scope.

\subparagraph{Variables}
\begin{itemize}
\item \verb+SYSCTL_STRING+
\item \verb+SYSCTL_INT+
\item \verb+SYSCTL_UINT+
\item \verb+SYSCTL_LONG+
\item \verb+SYSCTL_ULONG+
\item \verb+SYSCTL_COUNTER_U64+
\end{itemize}

\subparagraph{Procedures}
\lstinputlisting[label=list:sysctl_proc,caption=Adding a sysctl prodedure.]{sys/sysctl_proc.c}

\subparagraph{Feature test variables}
\verb+FEATURE(name, desc)+
